---
title: "COI_taxonomy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{COI_taxonomy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we will construct a taxonmic traing "database" for
the Cytochrome C oxidase (the "Folmer region" of it to be accurate,
which is the region most commonly sequenced as a "DNA barcode").

```{r setup}
library(taxAnyEukAmp)
```

After loading the package we count (query the database for their
numer) and download the sequences. As COI is a protein coding marker
we have to query the respective data ("coding release"). We read the
resulting downloaded files into a DNAStringSet and check whether the
number of seequences in the download fits with our expectation from
the database query.

The remaining processing is similar to non-coding (rRNA) sequences.

```{r download-clean}
COIDownloads <- getENAdownloads("COX1", "/SAN/db/ENA_marker/COI/", data = "coding_release")
COISeq <- Biostrings::readDNAStringSet(getFiles(COIDownloads))
length(COISeq) - getShouldBe(COIDownloads)

nonACGT <- !isACGT(as.character(COISeq))

accessions <- getAccession4ENAname(names(COISeq))
TaxIDs <- taxonomizr::accessionToTaxa(accessions, "/SAN/db/taxonomy/taxonomizr.sql")
duplicates <- duplicated(COISeq)&duplicated(TaxIDs)           

table(is.na(TaxIDs))
table(duplicates)
```



```{r taxon-clean}
taxonomy <- taxonomizr::getTaxonomy(TaxIDs, "/SAN/db/taxonomy/taxonomizr.sql")

## we can store this conveniently in one object to work with without
## keeping track of sequences and taxnomy seperately
COISeqtaxed <- createTaxedSeq(COISeq, taxonomy)

         
badTaxa <- getBadTaxa(COISeqtaxed, fromN=2)
badSpecies <- getBadSpecies(COISeqtaxed)         
table(badTaxa, badSpecies)
```

```{r align-clean}
COIClean <- subselectTaxedSeq(COISeqtaxed,
                             !nonACGT & !duplicates & !badTaxa & !badSpecies &
                             width(COISeq)>600) ## this excludes below 600 bases



COIMatrices <- getMatrices(COIClean, mc.cores=20)

outliers <- getOutliers(COIMatrices, COIClean)
subsequences <- getSubsequeces(COIMatrices, COIClean, mc.cores=20)

table(outliers, subsequences)

COICurated <- subselectTaxedSeq(COIClean, !outliers & !subsequences)
```

```{r train-clean}
IdTaxaResultsCOI <- idTaxaTrainAndClean(COICurated, 
                                        taxonomyCurated,
                                        badRetain=100)

writeFasta(IdTaxaResultsCOI[[1]], "/SAN/db/blastdb/COI_ENA/Length_600.fasta")

saveRDS(IdTaxaResultsCOI[[3]], "/SAN/db/IdTaxa/COI_Lenght600.rds")

```


